A sheaf--function dictionary for algebraic tori over local fields

### 2013-10-10 shanghai.tex
Massaged the new paragraph in Section 6.

We fix a prime $\ell$, invertible in $\Fq$.
We will work with constructible $\ell$-adic sheaves \cite{deligne:80a}*{\S 1.1}, \cite{SGA5}*{Expos\'es V, VI} 
on $\QC$-schemes, employing the standard formalism.
%see, in particular, \cite{SGA5}*{Expos\'e VI,\S 2}, where pullback and tensor products are developed.
We also make extensive use of the external tensor product of $\ell$-adic sheaves, defined as follows: if $\mathcal{F}$ and $\mathcal{G}$ are constructible $\ell$-adic sheaves on schemes $X$ and $Y$ and $p_X : X\times Y\to X$ and $p_Y : X\times Y \to Y$ are the projections, then $\mathcal{F}\boxtimes \mathcal{G} \ceq p_X^* \mathcal{F} \otimes p_Y^*\mathcal{G}$.

@@

To prove the next proposition, we will need a certain perspective on local systems.
 If $\qcs{E}$ is any local system on a connected scheme $G$ and $\bg$ is a geometric point on $G$,
 we may associate to $\qcs{E}$ a character of the \'etale fundamental group $\pi_1(G, \bg)$ \David{add citation}.\Clifton{I'll find one.}
 Since $\pi_1(G, \bg)$ is defined as a projective limit of automorphism groups of finite \'etale covers, a general
 local system will correspond to a projective limit of characters of such groups.  But some local systems can be
 described in terms of a single cover $f: H \to G$ and a representation $\psi: \ker f \to \Aut(V)$;
 two such descriptions yield isomorphic local systems if and only if they are related by a weak
 isomorphism in $C(G)$.  While the description of local systems in terms of $\pi_1(G, \bg)$ does
 not generalize well to non-connected $G$, the description in terms of $C(G)$ does.  \David{Look okay?}

@@

Now we localize $C(G)$.
Let us say that $(g,T)$ in $C(G)$ is a \emph{weak isomorphism} if $g$ is surjective
and $T$ is an isomorphism of vector spaces.  Recall from
\cite{kashiwara-schapira:CatgoriesSheaves}*{Def 7.1.5} that a \emph{left multiplicative system}
is a family of morphisms $W$ that
\begin{enumerate}
\item contains all isomorphisms,
\item is closed under composition,
\item is closed under pullback by arbitrary morphisms in $C(G)$,
\item and if the cokernel \David{Is $C(G)$ abelian?} \Clifton{Yes. In fact, that's stated, implicitly, in the paragraph above: a tensor category is an abelian category.} of a morphism $\phi \in \mor C(G)$ is in
$W$ then so is the kernel of $\phi$.
\end{enumerate}

@@

\begin{lemma}
The family of weak isomorphisms in $C(G)$ is a saturated, left multiplicative system.
\end{lemma}
\begin{proof}
Any isomorphism $(g, T)$ in $C(G)$ has both $g$ and $T$ isomorphisms, and thus is a weak isomorphism.
Clearly, the composition of two weak isomorphisms is a weak isomorphism.  The remaining properties
follow from the fact that in a weak isomorphism $(g, T)$, $g$ is an \'etale cover, together with standard
facts on \'etale covers and pullbacks. \David{Is this sufficient?}\Clifton{Yes, I think so.}
\end{proof}
Let $W$ be the family of weak isomorphisms and \cdef{$C(G)[W^{-1}]$} be the localization of $C(G)$ at $W$.
A morphism in $C(G)[W^{-1}]$ is an equivalence class of diagrams
%  \begin{equation}\label{eq:morGCf}
\[   
\begin{tikzcd}[column sep=30]
   (f,\psi) \arrow{r}{(g,T)} & (f',\psi') & \arrow[swap]{l}{(h,U)} (f_1,\psi_1)
   \end{tikzcd}
\]
%  \end{equation}
 where $(g,T) : (f,\psi) \to (f',\psi')$ is any map in $C(G)$ 
 and $(h,U) :  (f',\psi') \to (f_1,\psi_1)$ is a weak isomorphism in $C(G)$ \cite{kashiwara-schapira:CatgoriesSheaves}*{Thm 7.1.16}.
%Likewise, $C(G)[W^{-1}]$ is a tensor category.

>>>>

We fix a prime $\ell$, invertible in $\Fq$.
We will work with constructible $\ell$-adic sheaves \citelist{\cite{deligne:80a}*{\S 1.1} \cite{SGA5}*{Expos\'es V, VI}}
on $\QC$-schemes, employing the standard formalism.
%see, in particular, \cite{SGA5}*{Expos\'e VI,\S 2}, where pullback and tensor products are developed.
We also make extensive use of the external tensor product of $\ell$-adic sheaves, defined as follows: if $\mathcal{F}$ and $\mathcal{G}$ are constructible $\ell$-adic sheaves on schemes $X$ and $Y$ and $p_X : X\times Y\to X$ and $p_Y : X\times Y \to Y$ are the projections, then $\mathcal{F}\boxtimes \mathcal{G} \ceq p_X^* \mathcal{F} \otimes p_Y^*\mathcal{G}$.

@@

The notion of a finite quasicharacter sheaf is informed 
by a certain perspective on local systems on connected groups.
Recall that every geometric point $\bg$ on a connected scheme $G$ 
determines an equivalence between the categories of $\ell$-adic local systems on $G$ 
and $\ell$-adic representations of the \'etale fundamental group $\pi_1(G, \bg)$ 
\citelist{\cite{SGA5}*{VI,Prop. 1.2.5} \cite{deligne:80a}*{1.1.1}}.
 Since $\pi_1(G, \bg)$ is a projective limit of automorphism groups of finite \'etale covers, 
 a general local system corresponds to a projective limit of characters of such groups.  
 But some local systems can be described in terms of a single cover $f: H \to G$ 
 and a representation $\psi: \ker f \to \Aut(V)$;
 two such descriptions yield isomorphic local systems if and only if they are related by a weak
 isomorphism in $C(G)$.  While the description of local systems in terms of $\pi_1(G, \bg)$ does
 not generalize well to non-connected $G$, a perspective using finite \'etale covers does.

@@

Now we localize $C(G)$.
Let us say that $(g,T)$ in $C(G)$ is a \emph{weak isomorphism} if $g$ is surjective
and $T$ is an isomorphism of vector spaces.  Recall from
\cite{kashiwara-schapira:CatgoriesSheaves}*{Def 7.1.5} that a \emph{left multiplicative system}
is a family of morphisms $W$ that
\begin{enumerate}
\item contains all isomorphisms,
\item is closed under composition,
\item is closed under pullback by arbitrary morphisms in $C(G)$,
\item and if the cokernel of a morphism $\phi \in \mor C(G)$ is in
$W$ then so is the kernel of $\phi$.
\end{enumerate}

@@

\begin{lemma}
The family of weak isomorphisms in $C(G)$ is a saturated, left multiplicative system.
\end{lemma}
\begin{proof}
Any isomorphism $(g, T)$ in $C(G)$ has both $g$ and $T$ isomorphisms, and thus is a weak isomorphism.
Clearly, the composition of two weak isomorphisms is a weak isomorphism.  The remaining properties
follow from the fact that in a weak isomorphism $(g, T)$, $g$ is an \'etale cover, together with standard
facts on \'etale covers and pullbacks.
\end{proof}
Let $W$ be the family of weak isomorphisms and \cdef{$C(G)[W^{-1}]$} be the localization of $C(G)$ at $W$.
A morphism in $C(G)[W^{-1}]$ is an equivalence class of diagrams
%  \begin{equation}\label{eq:morGCf}
\[   
\begin{tikzcd}[column sep=30]
   (f,\psi) \arrow{r}{(g,T)} & (f',\psi') & \arrow[swap]{l}{(h,U)} (f_1,\psi_1)
   \end{tikzcd}
\]
%  \end{equation}
 where $(g,T) : (f,\psi) \to (f',\psi')$ is any map in $C(G)$ 
 and $(h,U) :  (f',\psi') \to (f_1,\psi_1)$ is a weak isomorphism in $C(G)$ \cite{kashiwara-schapira:CatgoriesSheaves}*{Thm 7.1.16}.
%Likewise, $C(G)[W^{-1}]$ is a tensor category.
